---
title: "Buoy-Project"
author: "Ruxin Liu"
date: "9/24/2020"
output: pdf_document
---

```{r}
library(tidyverse)
library(stringr)

```
# Importing the data 

```{r}
# The first part of the data URL
url_1 <- "http://www.ndbc.noaa.gov/view_text_file.php?filename=mlrf1h"
# The second part of the data URL
url_2 <- ".txt.gz&dir=data/historical/stdmet/"
# Read the latest 20 years of data
years <- c(2000:2019)
# The complete 20 URLs
urls <- str_c(url_1, years, url_2, sep = "")
filenames <- str_c("mr", years, sep = "")

```


## Since the buoy data was collected over years, there were updates and changes on how many variables are measured. During the exploring process, it is found that the data from the year 2000 to 2004 only contained 17 columns, while starting from the year 2005, the data set had one extra column mm, which is the minute variable. Since the mm variable only contained values of 0 and the earlier years didn't have this information available, the mm variable will not be used in the analysis and can be removed from the data.


```{r}
# Load in the 20-year data from the NDBC website
# Year 2000 - 2006
for(i in 1:7){
  suppressMessages(
    # Fill any missing values with NA:
    assign(filenames[i], read.table(urls[i], header = TRUE, fill = TRUE))
  )
  
}

# Year 2007 - 2019: Column names started with #, which will cause issue
for(i in 8:20){
  suppressMessages(
    # Fill any missing values with NA and use the same column names as year 2006
    assign(filenames[i], read.table(urls[i], header = FALSE, 
                                    fill = TRUE, col.names = colnames(mr2006))),
   
  )

}

```

```{r}

```

# Cleaning the data 


```{r}

```


